--- 
title: "ICT Gender-Equality Paradox"
author: "Mandy Davis"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [ict-gep.bib]
citation-style: apa.csl
link-citations: yes
github-repo: mandydavis/ict-gep-re-analysis
description: "An overview and tour of the steps taken to re-analyze the ICT Gender-Equality Paradox."
---

```{r setup, include=FALSE}
# packages needed for both SAPA.r and thesis stats.R
library(readxl)
library(ggplot2)
library(psych)
library(psychTools)
library(reshape2)
library(janitor)
library(tidyverse)
library(ggrepel)
library(directlabels)
library(ggforce)
library(gridExtra)
library(gtable)
library(dplyr)
library(forcats)
require(devtools)

# necessary mapping packages
theme_set(theme_bw())
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library("rgeos")
library(magrittr) # pipes
library(lintr) # code linting
library(raster) # raster handling (needed for relief)
library(viridis) # viridis color scale
library('dichromat')
```

```{r themes, include = FALSE}
# set theme for maps & pies
# set defaults
default_font_family <- 'Tahoma'
default_font_color <- '#22211d'
default_background_color <- '#f5f5f2'
theme_map <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = default_font_family,
                          color = default_font_color),
      # remove all axes
      axis.line = element_blank(),
      # axis.text.x = element_blank(),
      # axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      # add a subtle grid
      # panel.grid.major = element_line(color = "#dbdbd9", size = 0.2),
      # panel.grid.minor = element_blank(),
      # background colors
      plot.background = element_rect(fill = default_background_color,
                                     color = NA),
      panel.background = element_rect(fill = default_background_color,
                                      color = NA),
      legend.background = element_rect(fill = default_background_color,
                                       color = NA),
      # borders and margins
      plot.margin = unit(c(.5, .5, .2, .5), "cm"),
      panel.border = element_blank(),
      panel.spacing = unit(c(-.1, 0.2, .2, 0.2), "cm"),
      # titles
      legend.title = element_text(size = 9, margin = margin(-.2,0,.2,0, unit = 'cm')),
      legend.text = element_text(size = 9, hjust = 0,
                                 color = default_font_color),
      legend.position="bottom", 
      legend.box = "horizontal" ,
      plot.title = element_text(size = 14, hjust = 0.5,
                                color = default_font_color),
      plot.subtitle = element_text(size = 9, hjust = 0.5,
                                   color = default_font_color,
                                   margin = margin(b = .5,
                                                   t = 0,
                                                   l = 2,
                                                   unit = "cm"),
                                   debug = F),
      # captions
      plot.caption = element_text(size = 7,
                                  hjust = .5,
                                  margin = margin(t = 0.2,
                                                  b = 0,
                                                  unit = "cm"),
                                  color = "#939184"),
      ...
    )
}
theme_pie <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = default_font_family,
                          color = default_font_color),
      # remove all axes
      axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      # add a subtle grid
      # panel.grid.major = element_line(color = "#dbdbd9", size = 0.2),
      panel.grid.minor = element_blank(),
      # background colors
      plot.background = element_rect(fill = default_background_color,
                                     color = NA),
      panel.background = element_rect(fill = default_background_color,
                                      color = NA),
      legend.background = element_rect(fill = default_background_color,
                                       color = NA),
      # borders and margins
      plot.margin = unit(c(.5, .5, .2, .5), "cm"),
      panel.border = element_blank(),
      panel.spacing = unit(c(-.1, 0.2, .2, 0.2), "cm"),
      # titles
      legend.title = element_blank(),
      legend.text = element_text(size = 9, hjust = 0,
                                 color = default_font_color),
      legend.position="top", 
      legend.box = "horizontal" ,
      plot.title = element_text(size = 9, hjust = 0.5,
                                color = default_font_color),
      plot.subtitle = element_text(size = 9, hjust = 0.5,
                                   color = default_font_color,
                                   margin = margin(b = .5,
                                                   t = 0,
                                                   l = 2,
                                                   unit = "cm"),
                                   debug = F),
      # captions
      plot.caption = element_text(size = 7,
                                  hjust = .5,
                                  margin = margin(t = 0.2,
                                                  b = 0,
                                                  unit = "cm"),
                                  color = "#939184"),
      ...
    )
}
```

```{r SAPA, include=FALSE}
load("./data & analysis/data/SAPA/SAPA.rdata")
# rename the monster
SAPA <- SAPAdata18aug2010thru7feb2017
rm(SAPAdata18aug2010thru7feb2017)
# reset ItemLists with the file Bill provided (running this line will replace the existing one from the earlier load)
load("./data & analysis/data/SAPA/ItemLists.rda")

# country/participant count
countries_tibble <- SAPA %>%
  mutate(country = forcats::fct_explicit_na(country)) %>%
  group_by(country) %>%
  summarize(N = n()) %>%
  arrange(desc(N))

# countries with >= 500 participants
countries_tibble_500 <- filter(countries_tibble, N >= 500)

# a new version of the SAPA data, which includes only those participants who are from countries that have >= 500 total participants
SAPA_500 <- subset(SAPA, SAPA$country %in% countries_tibble_500$country)

# merge SAPA_500 with GGGI data (we have SAPA data for 2010-2017, so select same years for GGGI)
# update: there is no 2017 in the GGGI data (?), so the last year I include is 2016, as the SAPA data only goes through Feb 2017.
ggi_raw <- read.csv("./data & analysis/data/ggi.csv", header = TRUE)
# reduces the data frame to contain only the relevant data, aka the ggi index value
ggi_index <- subset(ggi_raw, Indicator=='Overall Global Gender Gap Index' & Subindicator.Type== 'Index')
# find average GGI from 2013-2018
ggi_index$avg_ggi <- rowMeans(ggi_index[,10:16], na.rm = TRUE)
```

# Exploratory Data Analysis

## ICT Graduation Trends (Country Comparison)
```{r ICT_pies, echo=FALSE, warning=FALSE, message=FALSE}
####### visualize degree popularity (what % of all women and men tertiary graduates studied ICT?):
percent_of_women <- read.csv("./data & analysis/data/percent_of_women.csv")
percent_of_men <- read.csv("./data & analysis/data/percent_of_men.csv")

# munge data [** keep in mind this is averaging across years **]
percent_of_women <- subset(percent_of_women, Indicator == 'Percentage of female graduates from tertiary education graduating from Information and Communication Technologies programmes, female (%)')
percent_of_women <- aggregate(percent_of_women[,'Value'], list(percent_of_women$LOCATION), mean)
colnames(percent_of_women) <- c("iso_a3", "percent_of_women")

# do the same for the men
percent_of_men <- subset(percent_of_men, Indicator == 'Percentage of male graduates from tertiary education graduating from Information and Communication Technologies programmes, male (%)')
percent_of_men <- aggregate(percent_of_men[,'Value'], list(percent_of_men$LOCATION), mean)
colnames(percent_of_men) <- c("iso_a3", "percent_of_men")

# combine the percent_of_women and percent_of_men tables
percent_of_gender <- inner_join(percent_of_women, percent_of_men, by = "iso_a3")

# load ICT data
ict_raw <- read.csv("./data & analysis/data/ict.csv")
# the distinct country codes in case you need them
ict_countries <- levels(ict_raw[,'LOCATION'])
# get average across all years that each country has data for (2013-2019)
ict_averages <- aggregate(ict_raw[,'Value'], list(ict_raw$LOCATION), mean)
# rename columns
colnames(ict_averages) <- c('iso_a3', 'average_ICT_w')

# ---------------------------
# Qatar 
# ---------------------------

average_ICT_w <- subset(ict_averages, subset = ict_averages["iso_a3"] == "QAT")$average_ICT_w
QAT_percent_ICT <- data.frame("gender" = c("Women", "Men"), "percent" = c(average_ICT_w, 100 - average_ICT_w))

# (load percent_of_gender data) create tables that will work well to construct percent_of_m and percent_of_w pie charts
QAT_percent_w <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "QAT")$percent_of_women
QAT_percent_of_w = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(QAT_percent_w, 100 - QAT_percent_w))
QAT_percent_m <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "QAT")$percent_of_men
QAT_percent_of_m = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(QAT_percent_m, 100 - QAT_percent_m))

# create pie chart (% of ICT graduates who are women)
QAT_p_of_ICT <- ggplot(QAT_percent_ICT, aes(x = "", y = percent, fill = gender)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#1FBFC3", "#F57670")) +
  labs(x= NULL, y = NULL, title = "% of ICT tertiary graduates who are..") +
  theme_pie()

# Qatar percent of women
QAT_p_of_women_plot <- ggplot(QAT_percent_of_w, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of women who pursued...") +
  theme_pie()

# Qatar percent of men
QAT_p_of_men_plot <- ggplot(QAT_percent_of_m, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of men who pursued...") +
  theme_pie() 

# ---------------------------
# Morocco 
# ---------------------------

average_ICT_w <- subset(ict_averages, subset = ict_averages["iso_a3"] == "MAR")$average_ICT_w
MAR_percent_ICT <- data.frame("gender" = c("Women", "Men"), "percent" = c(average_ICT_w, 100 - average_ICT_w))

MAR_percent_w <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "MAR")$percent_of_women
MAR_percent_of_w = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(MAR_percent_w, 100 - MAR_percent_w))
MAR_percent_m <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "MAR")$percent_of_men
MAR_percent_of_m = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(MAR_percent_m, 100 - MAR_percent_m))

# create pie chart (% of ICT graduates who are women)
MAR_p_of_ICT <- ggplot(MAR_percent_ICT, aes(x = "", y = percent, fill = gender)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#1FBFC3", "#F57670")) +
  labs(x= NULL, y = NULL, title = "% of ICT tertiary graduates who are..") +
  theme_pie()

# Qatar percent of women
MAR_p_of_women_plot <- ggplot(MAR_percent_of_w, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of women who pursued...") +
  theme_pie()

# Qatar percent of men
MAR_p_of_men_plot <- ggplot(MAR_percent_of_m, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of men who pursued...") +
  theme_pie() 

# ---------------------------
# United Kingdom 
# ---------------------------

average_ICT_w <- subset(ict_averages, subset = ict_averages["iso_a3"] == "GBR")$average_ICT_w
GBR_percent_ICT <- data.frame("gender" = c("Women", "Men"), "percent" = c(average_ICT_w, 100 - average_ICT_w))

GBR_percent_w <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "GBR")$percent_of_women
GBR_percent_of_w = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(GBR_percent_w, 100 - GBR_percent_w))
GBR_percent_m <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "GBR")$percent_of_men
GBR_percent_of_m = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(GBR_percent_m, 100 - GBR_percent_m))

# create pie chart (% of ICT graduates who are women)
GBR_p_of_ICT <- ggplot(GBR_percent_ICT, aes(x = "", y = percent, fill = gender)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#1FBFC3", "#F57670")) +
  labs(x= NULL, y = NULL, title = "% of ICT tertiary graduates who are..") +
  theme_pie()

# Qatar percent of women
GBR_p_of_women_plot <- ggplot(GBR_percent_of_w, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of women who pursued...") +
  theme_pie()

# Qatar percent of men
GBR_p_of_men_plot <- ggplot(GBR_percent_of_m, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of men who pursued...") +
  theme_pie() 


```

### Qatar
```{r qatar_pie, echo=FALSE, warning=FALSE, message=FALSE}
grid.arrange(QAT_p_of_ICT, QAT_p_of_women_plot, QAT_p_of_men_plot, nrow = 1)
```

### Morocco
```{r morocco_pie, echo=FALSE, warning=FALSE, message=FALSE}
grid.arrange(MAR_p_of_ICT, MAR_p_of_women_plot, MAR_p_of_men_plot, nrow = 1)
```

### United Kingdom
```{r uk_pie, echo=FALSE, warning=FALSE, message=FALSE}
grid.arrange(GBR_p_of_ICT, GBR_p_of_women_plot, GBR_p_of_men_plot, nrow = 1)
```

## ICT Graduation Trends (Change Over Time)
```{r ICT_change, echo=FALSE, warning=FALSE}
###    change over time    ###
country_codes <- c("QAT","MAR","SGP","EST","NZL","GBR","FRA")
ICT_circ <- subset(ict_raw, subset = ict_raw$LOCATION %in% country_codes)
# ICT_circ_df <- data.frame(row.names = c("QAT","MAR","SGP","EST","NZL","GBR","FRA"))
# ICT_circ_df <- t(ICT_circ_df)
ICT_circ_df <- data.frame(row.names = c(2013:2018))

ICT_circ_df$QAT <- NA
ICT_circ_df$MAR <- NA
ICT_circ_df$SGP <- NA
ICT_circ_df$EST <- NA
ICT_circ_df$NZL <- NA
ICT_circ_df$GBR <- NA
ICT_circ_df$FRA <- NA

for (country_code in colnames(ICT_circ_df)) {
  country_years <- subset(ICT_circ$TIME, ICT_circ$LOCATION == country_code)
  for (year in strtoi(rownames(ICT_circ_df))) {
    if(year %in% country_years) {
      country_rows <- subset(ICT_circ, subset = ICT_circ$LOCATION==country_code)
      ICT_circ_df[year,country_code] <- subset(country_rows, subset = country_rows$TIME==year)$Value
    }
  }
}

# chop off the rest of the rows
ICT_circ_df <- ICT_circ_df[2013:2018,]
rownames(ICT_circ_df) <- c(2013:2018)

ICT_circ_df$year <- as.numeric(row.names(ICT_circ_df))



# plot ICT (women) graduation rates usage for my 7 countries using the years available:
ICT_change <- ggplot(ICT_circ_df, aes(x = year)) +
  # ylim(0, 100) +
  geom_line(aes(y = as.numeric(as.character(MAR))), group = 1, color = 'red') + 
  geom_line(aes(y = as.numeric(as.character(EST))), group = 1, color = 'blue') +
  geom_line(aes(y = as.numeric(as.character(FRA))), group = 1, color = 'pink') +
  geom_line(aes(y = as.numeric(as.character(QAT))), group = 1, color = 'green') +
  geom_line(aes(y = as.numeric(as.character(GBR))), group = 1, color = 'gray') +
  geom_line(aes(y = as.numeric(as.character(NZL))), group = 1, color = 'orange') +
  geom_line(aes(y = as.numeric(as.character(SGP))), group = 1, color = 'black') +
  labs(title = "Change in the percent of ICT tertiary graduates who are women over time", subtitle = "subtitle goes here", y= "percent of ICT tertiary graduates who are women") +
  geom_text(x= 2016.25, y = 16.45824, label = "France", size = 3, color= "pink") +
  geom_text(x= 2017.25, y = 41.28349, label = "Morocco", size = 3, color= "red") +
  geom_text(x= 2017.3, y = 32.22160, label = "Singapore", size = 3, color= "black") +
  geom_text(x= 2016.15, y = 19.38466, label = "UK", size = 3, color= "gray") +
  geom_text(x= 2017.35, y = 23.10945, label = "New Zealand", size = 3, color= "orange") +
  geom_text(x= 2017.25, y = 28.89201, label = "Estonia", size = 3, color= "blue") +
  geom_text(x= 2018, y = 54.2, label = "Qatar", size = 3, color= "green") +
  theme_map()

ICT_change

```

## Internet Usage

```{r internet, echo=FALSE, warning=FALSE}
internet_usage <- read_excel("./data & analysis/data/Internet Usage downloaded 08.24.20 (World Bank)/internet_usage.xls")

# useful way to rename one specific column
colnames(internet_usage)[1] <- "country_name"

# countries of interest
countries <- cs(Morocco, "United Kingdom", France, Estonia, Qatar, Singapore, "New Zealand")

internet_usage_selection <- as.data.frame(t(subset(internet_usage, country_name %in% countries)[, -c(2, 3, 4)]))
# make country names the header
internet_usage_selection <- row_to_names(internet_usage_selection, row_number = 1) 
# make years (currently as rownames) to their own column
internet_usage_selection$year <- as.numeric(row.names(internet_usage_selection))
# we don't care about the years before the internet existed
internet_usage_selection <- subset(internet_usage_selection, subset = year > 1989)

# internet <- internet_usage_selection %>%
#   select(date, psavert, uempmed) %>%
#   gather(key = "variable", value = "value", -date)


# plot internet usage for my 7 countries from 1990-2020:
internet <- ggplot(internet_usage_selection, aes(x = year)) +
  # ylim(0, 100) +
  geom_line(aes(y = as.numeric(as.character(Morocco))), group = 1, color = 'red') + 
  geom_line(aes(y = as.numeric(as.character(Estonia))), group = 1, color = 'blue') +
  geom_line(aes(y = as.numeric(as.character(France))), group = 1, color = 'pink') +
  geom_line(aes(y = as.numeric(as.character(Qatar))), group = 1, color = 'green') +
  geom_line(aes(y = as.numeric(as.character(`United Kingdom`))), group = 1, color = 'gray') +
  geom_line(aes(y = as.numeric(as.character(`New Zealand`))), group = 1, color = 'orange') +
  geom_line(aes(y = as.numeric(as.character(Singapore))), group = 1, color = 'black') +
  labs(title = "Individuals using the Internet 1990-2018 (% of population)", subtitle = "Internet users are individuals who have used the Internet (from any location) in the last 3 months.\n The Internet can be used via a computer, mobile phone, personal digital assistant, games machine, digital TV etc.", y = "Individuals using the Internet (% of population)", x = "Year", caption = "Source: World Bank 2020, Data: International Telecommunication Union, World Telecommunication/ICT Development Report and database") +
  geom_text(x= 2019, y = 82.04319, label = "France", size = 3, color= "pink") +
  geom_text(x= 2019, y = 64.80387, label = "Morocco", size = 3, color= "red") +
  geom_text(x= 2019, y = 88.16564, label = "Singapore", size = 3, color= "black") +
  geom_text(x= 2019, y = 94.89674, label = "UK", size = 3, color= "gray") +
  geom_text(x= 2018, y = 90.81109, label = "New Zealand", size = 3, color= "orange") +
  geom_text(x= 2019, y = 89.35701, label = "Estonia", size = 3, color= "blue") +
  geom_text(x= 2019, y = 99.65285, label = "Qatar", size = 3, color= "green") +
  theme_map()

internet
```

## Interest: Oregon Vocational Interest Scales (ORVIS) from the SAPA Project

The ORVIS contains seven total scales: Leadership, Organization, Altruism, Creativity, Analysis, Production, Adventure, and Erudation. The Analysis scale is the scale of interest for this research. Ten items compose this scale, all of which are positively scored. The items are as follows:

* Be a chemist
* Design a laboratory experiment
* Be a mathematician
* Explain scientific concepts to others
* Be a physicist
* Carry out medical research
* Be a scientific reporter
* Solve complex puzzles
* Develop a computer program
* Be a statistician

These data are a subet of the SAPA project. While the PISA dataset offer robust insights into the academic interests of fiften year-olds, less is known about adults' interests in these same areas. 

```{r descriptive_stats, include=FALSE}
# find gender break down
SAPA_500 %>%
  mutate(gender = forcats::fct_explicit_na(gender)) %>%
  group_by(gender) %>%
  summarize(N = n())

# find mean, median, and range of age and sd
median(SAPA_500$age) # 22
mean(SAPA_500$age) # 26
range(SAPA_500$age) # 14 - 90
```

Of the participants who selected a country that has 500 or more participants, 219,728 participants form the final dataset used for analysis. Participants choose between "male" and "female" to select their gender identity, with $63.52\%$ ($N = 139,567$) identifying as female and the other $36.48\%$ ($N = 80,162$) identifying as male. The median age of the particpants is 22, ranging from 14 to 90. 


