--- 
title: "ICT Gender-Equality Paradox"
author: "Mandy Davis"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [ict-gep.bib]
citation-style: apa.csl
link-citations: yes
github-repo: mandydavis/thesis
description: "Probing the ICT Gender-Equality Paradox."
---

```{r setup, include=FALSE}
# packages needed for both SAPA.r and thesis stats.R
library(readxl)
library(ggplot2)
library(psych)
library(psychTools)
library(reshape2)
library(janitor)
library(tidyverse)
library(ggrepel)
library(directlabels)
library(ggforce)
library(gridExtra)
library(gtable)
library(dplyr)
library(forcats)
require(devtools)
library(cowplot)
library(magick)

# necessary mapping packages
theme_set(theme_bw())
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library("rgeos")
library(magrittr) # pipes
library(lintr) # code linting
library(raster) # raster handling (needed for relief)
library(viridis) # viridis color scale
library('dichromat')
```

```{r themes, include = FALSE}
# set theme for maps & pies
# set defaults
default_font_family <- 'Tahoma'
default_font_color <- '#22211d'
default_background_color <- '#f5f5f2'
theme_map <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = default_font_family,
                          color = default_font_color),
      # remove all axes
      axis.line = element_blank(),
      # axis.text.x = element_blank(),
      # axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      # add a subtle grid
      # panel.grid.major = element_line(color = "#dbdbd9", size = 0.2),
      # panel.grid.minor = element_blank(),
      # background colors
      plot.background = element_rect(fill = default_background_color,
                                     color = NA),
      panel.background = element_rect(fill = default_background_color,
                                      color = NA),
      legend.background = element_rect(fill = default_background_color,
                                       color = NA),
      # borders and margins
      plot.margin = unit(c(.5, .5, .2, .5), "cm"),
      panel.border = element_blank(),
      panel.spacing = unit(c(-.1, 0.2, .2, 0.2), "cm"),
      # titles
      legend.title = element_text(size = 9, margin = margin(-.2,0,.2,0, unit = 'cm')),
      legend.text = element_text(size = 9, hjust = 0,
                                 color = default_font_color),
      legend.position="bottom", 
      legend.box = "horizontal" ,
      plot.title = element_text(size = 14, hjust = 0.5,
                                color = default_font_color),
      plot.subtitle = element_text(size = 9, hjust = 0.5,
                                   color = default_font_color,
                                   margin = margin(b = .5,
                                                   t = 0,
                                                   l = 2,
                                                   unit = "cm"),
                                   debug = F),
      # captions
      plot.caption = element_text(size = 7,
                                  hjust = .5,
                                  margin = margin(t = 0.2,
                                                  b = 0,
                                                  unit = "cm"),
                                  color = "#939184"),
      ...
    )
}
theme_pie <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = default_font_family,
                          color = default_font_color),
      # remove all axes
      axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      # add a subtle grid
      # panel.grid.major = element_line(color = "#dbdbd9", size = 0.2),
      panel.grid.minor = element_blank(),
      # background colors
      plot.background = element_rect(fill = "white",
                                     color = NA),
      panel.background = element_rect(fill = "white",
                                      color = NA),
      legend.background = element_rect(fill = "white",
                                       color = NA),
      # borders and margins
      plot.margin = unit(c(.5, .5, .2, .5), "cm"),
      panel.border = element_blank(),
      panel.spacing = unit(c(-.1, 0.2, .2, 0.2), "cm"),
      # titles
      legend.title = element_blank(),
      legend.text = element_text(size = 9, hjust = 0,
                                 color = default_font_color),
      legend.position="top", 
      legend.box = "horizontal" ,
      plot.title = element_text(size = 9, hjust = 0.5,
                                color = default_font_color),
      plot.subtitle = element_text(size = 9, hjust = 0.5,
                                   color = default_font_color,
                                   margin = margin(b = .5,
                                                   t = 0,
                                                   l = 2,
                                                   unit = "cm"),
                                   debug = F),
      # captions
      plot.caption = element_text(size = 7,
                                  hjust = .5,
                                  margin = margin(t = 0.2,
                                                  b = 0,
                                                  unit = "cm"),
                                  color = "#939184"),
      ...
    )
}
```

```{r SAPA, include=FALSE}
load("./data & analysis/data/SAPA/SAPA.rdata")
# rename the monster
SAPA <- SAPAdata18aug2010thru7feb2017
rm(SAPAdata18aug2010thru7feb2017)
# reset ItemLists with the file Bill provided (running this line will replace the existing one from the earlier load)
load("./data & analysis/data/SAPA/ItemLists.rda")

# country/participant count
countries_tibble <- SAPA %>%
  mutate(country = forcats::fct_explicit_na(country)) %>%
  group_by(country) %>%
  summarize(N = n()) %>%
  arrange(desc(N))

# countries with >= 500 participants
countries_tibble_500 <- filter(countries_tibble, N >= 500)

# a new version of the SAPA data, which includes only those participants who are from countries that have >= 500 total participants
SAPA_500 <- subset(SAPA, SAPA$country %in% countries_tibble_500$country)

# merge SAPA_500 with GGGI data (we have SAPA data for 2010-2017, so select same years for GGGI)
# update: there is no 2017 in the GGGI data (?), so the last year I include is 2016, as the SAPA data only goes through Feb 2017.
ggi_raw <- read.csv("./data & analysis/data/ggi.csv", header = TRUE)
# reduces the data frame to contain only the relevant data, aka the ggi index value
ggi_index <- subset(ggi_raw, Indicator=='Overall Global Gender Gap Index' & Subindicator.Type== 'Index')
# find average GGI from 2013-2018
ggi_index$avg_ggi <- rowMeans(ggi_index[,10:16], na.rm = TRUE)
```

# Exploratory Data Analysis

## The Paradox
Gender equality (as measured by the Global Gender Gap Index) is moderately, negatively correlated with the  percentage of women among ICT, adjusted for equal graduation numbers of men and women ($r = -0.47$, $p < 0.001$).

```{r paradox, echo=FALSE, message=FALSE, warning=FALSE}
# ict data 
# source:  http://data.uis.unesco.org/Index.aspx?DataSetCode=EDULIT_DS&popupcustomise=true&lang=en#
# path from source to data: education --> completion --> percentage of tertiary graduates --> percentage of female graduates by field of study
# note on country codes: to join the various data frames, it essential to include the iso_a3 coutnry codes in each of the original data frames,
# so be sure to select 'customise' --> 'table options' and finally, check the box to use codes for country
ict_raw <- read.csv("./data & analysis/data/ict.csv")
# the distinct country codes in case you need them
ict_countries <- levels(ict_raw[,'LOCATION'])
# get average across all years that each country has data for (2013-2019)
ict_averages <- aggregate(ict_raw[,'Value'], list(ict_raw$LOCATION), mean)
# rename columns
colnames(ict_averages) <- c('iso_a3', 'average_ict')

# percent_of_women data
percent_of_women <- read.csv("./data & analysis/data/percent_of_women.csv", header = TRUE)
percent_of_women <- subset(percent_of_women, Indicator == 'Percentage of female graduates from tertiary education graduating from Information and Communication Technologies programmes, female (%)')
percent_of_women_avg <- aggregate(percent_of_women[,'Value'], list(percent_of_women$LOCATION), mean)
# percent_of_men data
percent_of_men <- read.csv("./data & analysis/data/percent_of_men.csv", header = TRUE)
percent_of_men <- subset(percent_of_men, Indicator == 'Percentage of male graduates from tertiary education graduating from Information and Communication Technologies programmes, male (%)')
percent_of_men_avg <- aggregate(percent_of_men[,'Value'], list(percent_of_men$LOCATION), mean)
# add a/(a+b) column
percent_of_ict_adjusted <- percent_of_women_avg
percent_of_ict_adjusted$b <- percent_of_men_avg[,'x']
colnames(percent_of_ict_adjusted) <- cs(iso_a3, a, b)
# add a column for the percent_of_ict adjusted percentage
percent_of_ict_adjusted$adjusted <- 100*percent_of_ict_adjusted[,'a']/(percent_of_ict_adjusted[,'a']+percent_of_ict_adjusted[,'b'])
# add a column for the disparity percentage
percent_of_ict_adjusted$disparity <- percent_of_ict_adjusted[,'a'] - percent_of_ict_adjusted[,'b']


# percent_of_ict data + gggi data
ict_ggi <- merge(ict_averages, ggi_index, by.x='iso_a3', by.y='Country.ISO3')

# append the gggi data
ict_ggi_adjusted <- merge(percent_of_ict_adjusted, ggi_index, by.x='iso_a3', by.y='Country.ISO3')

# The formula used for calculating the propensity of women to graduate with STEM degrees was a /(a + b), 
# where a is the percentage of women who graduate with STEM [ICT] degrees (relative to all women graduating) and b is the percentage 
# of men who graduate with STEM [ICT] degrees (relative to all men graduating).

# so that we can label the dots with the actual country name
rownames(ict_ggi_adjusted) <- ict_ggi_adjusted[,'Country.Name']

# plot it
ggplot(ict_ggi_adjusted, aes(x=avg_ggi, y=adjusted, label = rownames(ict_ggi_adjusted))) + 
  geom_label(size = 2) +
  scale_color_distiller(palette = "YlGn", type = "seq", direction = -1) +
  geom_smooth(method=lm) +
  labs(x = 'Global Gender Gap Index (GGGI) (2013-2018)',
       y = 'Expected percent of women among ICT graduates, accounting for\n gender differences in overall graduation numbers',
       title = "ICT Gender Equality Paradox",
       subtitle = "Adjusting for gender differences in overall graduation numbers",
       caption = "Data: UNESCO Institute For Statistics, displaying the average of a country's available data from 2013-2019") +
  theme_map() + theme(axis.title=element_text(size=8), axis.text.x = element_text(), axis.text.y = element_text())
```

## ICT Graduation Rates in 7 Countries
**[in progress]** I am working on creating the visualization for all seven countries so that I can stack them and demonstrate the disconnect there can be between the overall ratio of men and women graduates and the *percentage of all men and all women* who pursue ICT.

```{r ICT_pies, echo=FALSE, warning=FALSE, message=FALSE}
####### visualize degree popularity (what % of all women and men tertiary graduates studied ICT?):
percent_of_women <- read.csv("./data & analysis/data/percent_of_women.csv")
percent_of_men <- read.csv("./data & analysis/data/percent_of_men.csv")

# munge data [** keep in mind this is averaging across years **]
percent_of_women <- subset(percent_of_women, Indicator == 'Percentage of female graduates from tertiary education graduating from Information and Communication Technologies programmes, female (%)')
percent_of_women <- aggregate(percent_of_women[,'Value'], list(percent_of_women$LOCATION), mean)
colnames(percent_of_women) <- c("iso_a3", "percent_of_women")

# do the same for the men
percent_of_men <- subset(percent_of_men, Indicator == 'Percentage of male graduates from tertiary education graduating from Information and Communication Technologies programmes, male (%)')
percent_of_men <- aggregate(percent_of_men[,'Value'], list(percent_of_men$LOCATION), mean)
colnames(percent_of_men) <- c("iso_a3", "percent_of_men")

# combine the percent_of_women and percent_of_men tables
percent_of_gender <- inner_join(percent_of_women, percent_of_men, by = "iso_a3")

# load ICT data
ict_raw <- read.csv("./data & analysis/data/ict.csv")
# the distinct country codes in case you need them
ict_countries <- levels(ict_raw[,'LOCATION'])
# get average across all years that each country has data for (2013-2019)
ict_averages <- aggregate(ict_raw[,'Value'], list(ict_raw$LOCATION), mean)
# rename columns
colnames(ict_averages) <- c('iso_a3', 'average_ICT_w')

# ---------------------------
# Qatar 
# ---------------------------

average_ICT_w <- subset(ict_averages, subset = ict_averages["iso_a3"] == "QAT")$average_ICT_w
QAT_percent_ICT <- data.frame("gender" = c("Women", "Men"), "percent" = c(average_ICT_w, 100 - average_ICT_w))

# (load percent_of_gender data) create tables that will work well to construct percent_of_m and percent_of_w pie charts
QAT_percent_w <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "QAT")$percent_of_women
QAT_percent_of_w = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(QAT_percent_w, 100 - QAT_percent_w))
QAT_percent_m <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "QAT")$percent_of_men
QAT_percent_of_m = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(QAT_percent_m, 100 - QAT_percent_m))

# create pie chart (% of ICT graduates who are women)
QAT_p_of_ICT <- ggplot(QAT_percent_ICT, aes(x = "", y = percent, fill = gender)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#1FBFC3", "#F57670")) +
  labs(x= NULL, y = NULL, title = "% of ICT graduates who are..") +
  theme_pie()

# Qatar percent of women
QAT_p_of_women_plot <- ggplot(QAT_percent_of_w, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of women who pursued...") +
  theme_pie()

# Qatar percent of men
QAT_p_of_men_plot <- ggplot(QAT_percent_of_m, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of men who pursued...") +
  theme_pie() 

# ---------------------------
# Morocco 
# ---------------------------

average_ICT_w <- subset(ict_averages, subset = ict_averages["iso_a3"] == "MAR")$average_ICT_w
MAR_percent_ICT <- data.frame("gender" = c("Women", "Men"), "percent" = c(average_ICT_w, 100 - average_ICT_w))

MAR_percent_w <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "MAR")$percent_of_women
MAR_percent_of_w = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(MAR_percent_w, 100 - MAR_percent_w))
MAR_percent_m <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "MAR")$percent_of_men
MAR_percent_of_m = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(MAR_percent_m, 100 - MAR_percent_m))

# create pie chart (% of ICT graduates who are women)
MAR_p_of_ICT <- ggplot(MAR_percent_ICT, aes(x = "", y = percent, fill = gender)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#1FBFC3", "#F57670")) +
  labs(x= NULL, y = NULL, title = "% of ICT tertiary graduates who are..") +
  theme_pie()

# Qatar percent of women
MAR_p_of_women_plot <- ggplot(MAR_percent_of_w, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of women who pursued...") +
  theme_pie()

# Qatar percent of men
MAR_p_of_men_plot <- ggplot(MAR_percent_of_m, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of men who pursued...") +
  theme_pie() 

# ---------------------------
# United Kingdom 
# ---------------------------

average_ICT_w <- subset(ict_averages, subset = ict_averages["iso_a3"] == "GBR")$average_ICT_w
GBR_percent_ICT <- data.frame("gender" = c("Women", "Men"), "percent" = c(average_ICT_w, 100 - average_ICT_w))

GBR_percent_w <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "GBR")$percent_of_women
GBR_percent_of_w = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(GBR_percent_w, 100 - GBR_percent_w))
GBR_percent_m <- subset(percent_of_gender, subset = percent_of_gender["iso_a3"] == "GBR")$percent_of_men
GBR_percent_of_m = data.frame("degree" = c("ICT", "Other Degree"), "percent" = c(GBR_percent_m, 100 - GBR_percent_m))

# create pie chart (% of ICT graduates who are women)
GBR_p_of_ICT <- ggplot(GBR_percent_ICT, aes(x = "", y = percent, fill = gender)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#1FBFC3", "#F57670")) +
  labs(x= NULL, y = NULL, title = "% of ICT tertiary graduates who are..") +
  theme_pie()

# Qatar percent of women
GBR_p_of_women_plot <- ggplot(GBR_percent_of_w, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of women who pursued...") +
  theme_pie()

# Qatar percent of men
GBR_p_of_men_plot <- ggplot(GBR_percent_of_m, aes(x = "", y = percent, fill = degree)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) + 
  geom_label(aes(label = paste((round(percent, digits = 1)), "%")), position = position_stack(vjust = 0.5), show.legend = FALSE, size = 5) +
  scale_fill_manual(values = c("#6874E8", "#9F9FED")) +
  labs(x= NULL, y = NULL, title = "% of men who pursued...") +
  theme_pie() 


```

```{r flagtest, echo=FALSE, warning=FALSE, message=FALSE}
qatar_flag <- ggdraw() + draw_image("./assets/qatar.png") + draw_label("Qatar", vjust = -2.5, fontfamily = "Tahoma", size = 26)
# grid.arrange(qatar_flag, QAT_p_of_ICT, nrow = 1)

grid.arrange(qatar_flag, QAT_p_of_ICT, QAT_p_of_women_plot, QAT_p_of_men_plot, nrow = 1)

```

<!-- ### grid.arrange test -->
<!-- ```{r tworows, echo=FALSE, warning=FALSE, message=FALSE} -->
<!-- grid.arrange(QAT_p_of_ICT, QAT_p_of_women_plot, QAT_p_of_men_plot, QAT_p_of_ICT, QAT_p_of_women_plot, QAT_p_of_men_plot, nrow = 2) -->
<!-- ``` -->


<!-- ### Qatar -->
<!-- ```{r qatar_pie, echo=FALSE, warning=FALSE, message=FALSE} -->
<!-- grid.arrange(QAT_p_of_ICT, QAT_p_of_women_plot, QAT_p_of_men_plot, nrow = 1) -->
<!-- ``` -->

<!-- ### Morocco -->
<!-- ```{r morocco_pie, echo=FALSE, warning=FALSE, message=FALSE} -->
<!-- grid.arrange(MAR_p_of_ICT, MAR_p_of_women_plot, MAR_p_of_men_plot, nrow = 1) -->
<!-- ``` -->

<!-- ### United Kingdom -->
<!-- ```{r uk_pie, echo=FALSE, warning=FALSE, message=FALSE} -->
<!-- grid.arrange(GBR_p_of_ICT, GBR_p_of_women_plot, GBR_p_of_men_plot, nrow = 1) -->
<!-- ``` -->

## ICT Graduation Trends (Change Over Time)
Less data than I would prefer, but that's helpful info too!

```{r ICT_change, echo=FALSE, warning=FALSE}
###    change over time    ###
country_codes <- c("QAT","MAR","SGP","EST","NZL","GBR","FRA")
ICT_circ <- subset(ict_raw, subset = ict_raw$LOCATION %in% country_codes)
# ICT_circ_df <- data.frame(row.names = c("QAT","MAR","SGP","EST","NZL","GBR","FRA"))
# ICT_circ_df <- t(ICT_circ_df)
ICT_circ_df <- data.frame(row.names = c(2013:2018))

ICT_circ_df$QAT <- NA
ICT_circ_df$MAR <- NA
ICT_circ_df$SGP <- NA
ICT_circ_df$EST <- NA
ICT_circ_df$NZL <- NA
ICT_circ_df$GBR <- NA
ICT_circ_df$FRA <- NA

for (country_code in colnames(ICT_circ_df)) {
  country_years <- subset(ICT_circ$TIME, ICT_circ$LOCATION == country_code)
  for (year in strtoi(rownames(ICT_circ_df))) {
    if(year %in% country_years) {
      country_rows <- subset(ICT_circ, subset = ICT_circ$LOCATION==country_code)
      ICT_circ_df[year,country_code] <- subset(country_rows, subset = country_rows$TIME==year)$Value
    }
  }
}

# chop off the rest of the rows
ICT_circ_df <- ICT_circ_df[2013:2018,]
rownames(ICT_circ_df) <- c(2013:2018)

ICT_circ_df$year <- as.numeric(row.names(ICT_circ_df))



# plot ICT (women) graduation rates usage for my 7 countries using the years available:
ICT_change <- ggplot(ICT_circ_df, aes(x = year)) +
  # ylim(0, 100) +
  geom_line(aes(y = as.numeric(as.character(MAR))), group = 1, color = 'red') + 
  geom_line(aes(y = as.numeric(as.character(EST))), group = 1, color = 'blue') +
  geom_line(aes(y = as.numeric(as.character(FRA))), group = 1, color = 'pink') +
  geom_line(aes(y = as.numeric(as.character(QAT))), group = 1, color = 'green') +
  geom_line(aes(y = as.numeric(as.character(GBR))), group = 1, color = 'gray') +
  geom_line(aes(y = as.numeric(as.character(NZL))), group = 1, color = 'orange') +
  geom_line(aes(y = as.numeric(as.character(SGP))), group = 1, color = 'black') +
  labs(title = "Change in the percent of ICT tertiary graduates who are women over time", y= "Percent of ICT tertiary graduates who are women") +
  geom_text(x= 2016.25, y = 16.45824, label = "France", size = 3, color= "pink") +
  geom_text(x= 2017.25, y = 41.28349, label = "Morocco", size = 3, color= "red") +
  geom_text(x= 2017.3, y = 32.22160, label = "Singapore", size = 3, color= "black") +
  geom_text(x= 2016.15, y = 19.38466, label = "UK", size = 3, color= "gray") +
  geom_text(x= 2017.35, y = 23.10945, label = "New Zealand", size = 3, color= "orange") +
  geom_text(x= 2017.25, y = 28.89201, label = "Estonia", size = 3, color= "blue") +
  geom_text(x= 2018, y = 54.2, label = "Qatar", size = 3, color= "green") +
  theme_map()

ICT_change

```

## Internet Usage
This is another reason I am so interested in comparing Qatar and Morocco.

```{r internet, echo=FALSE, warning=FALSE}
internet_usage <- read_excel("./data & analysis/data/Internet Usage downloaded 08.24.20 (World Bank)/internet_usage.xls")

# useful way to rename one specific column
colnames(internet_usage)[1] <- "country_name"

# countries of interest
countries <- cs(Morocco, "United Kingdom", France, Estonia, Qatar, Singapore, "New Zealand")

internet_usage_selection <- as.data.frame(t(subset(internet_usage, country_name %in% countries)[, -c(2, 3, 4)]))
# make country names the header
internet_usage_selection <- row_to_names(internet_usage_selection, row_number = 1) 
# make years (currently as rownames) to their own column
internet_usage_selection$year <- as.numeric(row.names(internet_usage_selection))
# we don't care about the years before the internet existed
internet_usage_selection <- subset(internet_usage_selection, subset = year > 1989)

# internet <- internet_usage_selection %>%
#   select(date, psavert, uempmed) %>%
#   gather(key = "variable", value = "value", -date)


# plot internet usage for my 7 countries from 1990-2020:
internet <- ggplot(internet_usage_selection, aes(x = year)) +
  # ylim(0, 100) +
  geom_line(aes(y = as.numeric(as.character(Morocco))), group = 1, color = 'red') + 
  geom_line(aes(y = as.numeric(as.character(Estonia))), group = 1, color = 'blue') +
  geom_line(aes(y = as.numeric(as.character(France))), group = 1, color = 'pink') +
  geom_line(aes(y = as.numeric(as.character(Qatar))), group = 1, color = 'green') +
  geom_line(aes(y = as.numeric(as.character(`United Kingdom`))), group = 1, color = 'gray') +
  geom_line(aes(y = as.numeric(as.character(`New Zealand`))), group = 1, color = 'orange') +
  geom_line(aes(y = as.numeric(as.character(Singapore))), group = 1, color = 'black') +
  labs(title = "Individuals using the Internet 1990-2018 (% of population)", subtitle = "Internet users are individuals who have used the Internet (from any location) in the last 3 months.\n The Internet can be used via a computer, mobile phone, personal digital assistant, games machine, digital TV etc.", y = "Individuals using the Internet (% of population)", x = "Year", caption = "Source: World Bank 2020, Data: International Telecommunication Union, World Telecommunication/ICT Development Report and database") +
  geom_text(x= 2019, y = 82.04319, label = "France", size = 3, color= "pink") +
  geom_text(x= 2019, y = 64.80387, label = "Morocco", size = 3, color= "red") +
  geom_text(x= 2019, y = 88.16564, label = "Singapore", size = 3, color= "black") +
  geom_text(x= 2019, y = 94.89674, label = "UK", size = 3, color= "gray") +
  geom_text(x= 2018, y = 90.81109, label = "New Zealand", size = 3, color= "orange") +
  geom_text(x= 2019, y = 89.35701, label = "Estonia", size = 3, color= "blue") +
  geom_text(x= 2019, y = 99.65285, label = "Qatar", size = 3, color= "green") +
  theme_map()

internet
```

## Interest: Oregon Vocational Interest Scales (ORVIS) from the SAPA Project

A common argument for why the STEM and ICT Gender-Equality paradoxes might exist surrounds the level of interest women have in the subjects. In countries with a high level of gender equality, some argue, women are free to pursue what they truly are interested in, even if those fields are less lucrative. 

In their controversial paper, Stoet & Geary examined this using the Programme for International Student Assessment (PISA) dataset. Interest is also important because of its relationship with achievement. The PISA data is valuable for understanding the interests of 15-year-olds who live in OECD countries, however, it lack insight on the interests of adults. 

The Synthetic Aperture Personality Assessment (SAPA) collects data that can fill that gap. SAPA is an online survey created by Northwestern psychologist William Revelle that anyone with access to a web browser and internet connection can take (https://www.sapa-project.org/). The Oregon Vocational Interest Scales (ORVIS) are a component of SAPA.

The ORVIS contains seven total scales: Leadership, Organization, Altruism, Creativity, Analysis, Production, Adventure, and Erudation. The Analysis scale is the scale of interest for this research. Ten items compose this scale, all of which are positively scored. The items are as follows:

* Be a chemist
* Design a laboratory experiment
* Be a mathematician
* Explain scientific concepts to others
* Be a physicist
* Carry out medical research
* Be a scientific reporter
* Solve complex puzzles
* Develop a computer program
* Be a statistician

```{r descriptive_stats, include=FALSE}
# find gender break down
SAPA_500 %>%
  mutate(gender = forcats::fct_explicit_na(gender)) %>%
  group_by(gender) %>%
  summarize(N = n())

# find mean, median, and range of age and sd
median(SAPA_500$age) # 22
mean(SAPA_500$age) # 26
range(SAPA_500$age) # 14 - 90
```

Of the people who took the SAPA test from August 18, 2010 through February 7, 2017, 219,728 indicated they are from a country that ended up having 500 or more subjects during that time period. These people form the final dataset used for analysis. Participants choose between "male" and "female" to select their gender identity, with $63.52\%$ ($N = 139,567$) identifying as female and the other $36.48\%$ ($N = 80,162$) identifying as male. The median age of the particpants is 22, ranging from 14 to 90. 

Below is a visualization of how these participants scored on their level of analytical interest.


```{r orvis_both, echo=FALSE, message=FALSE, warning=FALSE}
#### find the correlation between women's orvis_an scores and gggi ####
# 451:458 are the ORVIS scale keys
ORVIS_scores <- char2numeric(scoreItems(keys = ItemLists[451:458], items = SAPA_500, impute = "none")) # need the char2numeric?
ORVIS <- data.frame(ORVIS_scores$scores)
ORVIS_an <- data.frame(ORVIS$ORVIS_Analytical)
ORVIS_an$gender <- SAPA_500$gender
ORVIS_an$country <- SAPA_500$country

# filter data to just women
ORVIS_an_women <- subset(ORVIS_an, gender == "female")
colnames(ORVIS_an_women) <- c("ORVIS_an", "gender", "country")
# exclude NaN results
ORVIS_an_women <- subset(ORVIS_an_women, is.nan(ORVIS_an_women$ORVIS_an) == FALSE)
# 
# find average for each country? or is that bad bc then get rid of se?
ORVIS_an_women_avg <- aggregate(ORVIS_an_women[,'ORVIS_an'], list(ORVIS_an_women$country), mean)
colnames(ORVIS_an_women_avg) <- c("Country.ISO3", "ORVIS_an_avg")
# join the two dataframes
ORVIS_an_women_final <- inner_join(ORVIS_an_women_avg, ggi_index, by = "Country.ISO3")

ORVIS_an_men <- subset(ORVIS_an, gender == "male")
colnames(ORVIS_an_men) <- c("ORVIS_an", "gender", "country")
# exclude NaN results
ORVIS_an_men <- subset(ORVIS_an_men, is.nan(ORVIS_an_men$ORVIS_an) == FALSE)
# 
# find average for each country? or is that bad bc then get rid of se?
ORVIS_an_men_avg <- aggregate(ORVIS_an_men[,'ORVIS_an'], list(ORVIS_an_men$country), mean)
colnames(ORVIS_an_men_avg) <- c("Country.ISO3", "ORVIS_an_avg")
# join the two dataframes
ORVIS_an_men_final <- inner_join(ORVIS_an_men_avg, ggi_index, by = "Country.ISO3")

#### now, overlay the men's and women's correlations to view the opposite-ness ####
ORVIS_an_both <- ORVIS_an_men_final
ORVIS_an_both$ORVIS_an_women <- ORVIS_an_women_final$ORVIS_an_avg
# plot it
ggplot(ORVIS_an_both) +
  geom_smooth(aes(x = avg_ggi, y = ORVIS_an_avg), colour = "blue") + 
  geom_smooth(aes(x = avg_ggi, y = ORVIS_an_women), colour = "red") + 
  labs(title = "Relationship between men's and women's analytical interest and gender equality", 
       x = "Average GGGI, 2013-2016", y = "Average level of analytical interest, 2010-2017") +
  geom_text(x= 0.6, y = 3.5, label = "men", size = 3, color= "blue") +
  geom_text(x= 0.6, y = 2.85, label = "women", size = 3, color= "red", angle = 20) +
  theme_map()

```

Next up is the same graph, but only for women's interest and with country names included for reference.

```{r interest_country, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(ORVIS_an_men_final, aes(x = avg_ggi, y = ORVIS_an_avg)) +
  geom_smooth() + geom_label(aes(label = Country.Name)) + 
  labs(title = "Relationship between women's analytical interest and gender equality", 
       x = "Average GGGI, 2013-2016", y = "Average level of analytical interest, 2010-2017") +
  theme_map()
```


**What does this mean?** I don't know (yet). There seems to be no clear correlation like that of the ICT-GEP. However, it almost appears that men's and women's interest are inversely proportional, which is a really interesting lead, as men's and women's interest are theoretically independent and have no cap. Contrast this to the proportion of men and women there are in a given classroom—the proportions are dependent upon one another, so we would expect this relationship. In this situation, however, a relationship like this is not expected (without considering the broader context).

